---
title: "Graphs with Statistical Tests Attached"
author: "Barry DeCicco"
date: "2026-01-08"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

## Contents


This presentation covers some ways to run statistical test and to attached the output to graphs.   This replaces having to run the tests separately and to process the output until it is suitable for use.

My sources are the R Graph Gallery, LinkedIn posts and the Medium website.

- The package '**ggpubr**' (https://rpkgs.datanovia.com/ggpubr/)
- The package '**ggstatsplot**' (https://indrajeetpatil.github.io/ggstatsplot/, https://r-graph-gallery.com/package/ggstatsplot.html) 


## R Graph Gallery

This is a great resource (https://r-graph-gallery.com) which contains illustrations and code chunks for a large variety of plots, and variations within plot types.  It is the first place to check.


## Quick Start

You must have ggplot2 installed, then run the command:

``` {r  install_ggstatsplot}

install.packages("ggstatsplot")
library(ggstatsplot)


```


``` {r  quick_start}


library(ggstatsplot)
ggstatsplot::ggbetweenstats(data = mtcars, x = am, y = mpg, type = "p")

```

### Other Tests

There are nine functions, each for a specific statistical test.

#### Example: Scatterplot  (ggscatterstats)

This will create a scatterplot of the two variables, add marginal plots (histograms) 
and then add one test for the correlation and for the equality of means between the two
variables.


``` {r  ggscatterstats}


ggstatsplot::ggscatterstats(data = mtcars, x = hp, y = mpg, title = 'ggscatterstats')

```


#### Example: (ggwithinstats)

This looks at the data grouped by a single categorical variable.

**This example shows the integration of a ggstatsplot with other ggplot functions**

``` {r  ggwithinstats}

ggwithinstats(
  data = bugs_long,
  x = condition,
  y = desire,
  type = "nonparametric", ## type of statistical test
  xlab = "Condition", ## label for the x-axis
  ylab = "Desire to kill an artrhopod", ## label for the y-axis
  package = "yarrr", ## package from which color palette is to be taken
  palette = "info2", ## choosing a different color palette
  title = "ggwithinstats: Comparison of desire to kill bugs",
  caption = "Source: Ryan et al., 2013"
) + ## modifying the plot further
  ggplot2::scale_y_continuous(
    limits = c(0, 10),
    breaks = seq(from = 0, to = 10, by = 1)
  )
```

### Other ggstatsplot functions:

- ggbetweenstats, for comparisons between groups.
- ggwithinstats, for looking at distributions by level of a single categorical variable.
- gghistostats, for making a histogram and comparing if the mean is statistically significantly different from a specified value.
- ggcorrmat(), for making and examining correlalograms.
- ggpiestats, for making and testingp pie charts from contingency tables.
- ggbarstats, same as ggpiestats, but with bar charts.
- ggcoefstats, for dot-and-whisker plots, for regression coefficients and meta-analysis.






## References and Links:

- ggstatsplot:
  -  https://r-graph-gallery.com/package/ggstatsplot.html
  -  https://indrajeetpatil.github.io/ggstatsplot/
- R Graph Gallery:  https://r-graph-gallery.com/)



## What's Next?

I would like to explore AI *for* R/Rstudio and *within* R/Rstudio/Positron (more of these integrated AI applications.

My current plan is to present on AI *for* R/Rstudio in February, and
Integrated AI *within* R/Rstudio/Positron in March.



 
 
 

