---
title: "Graphs with Statistical Tests Attached"
author: "Barry DeCicco"
date: "2026-01-08"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Contents


This presentation covers some ways to run statistical test and to attached the output to graphs.   This replaces having to run the tests separately and to process the output until it is suitable for use.

My sources are the R Graph Gallery, LinkedIn posts and the Medium website.

- The package '**ggpubr**' (https://rpkgs.datanovia.com/ggpubr/)
- The package '**ggstatplot**' (https://indrajeetpatil.github.io/ggstatsplot/, https://r-graph-gallery.com/package/ggstatsplot.html) 


## R Graph Gallery

This is a great resource (https://r-graph-gallery.com) which contains illustrations and code chunks for a large variety of plots, and variations within plot types.  It is the first place to check.


## Quick Start

``` {r  quick_start}

library(ggstatsplot)
ggstatsplot::ggbetweenstats(data = mtcars, x = am, y = mpg, type = "p")

```

### Variations

### Bump Plot with highlighting.




## References and Links:

- ggside (https://cran.r-project.org/web/packages/ggpubr/index.html)
- ggstatsplot (https://ggrepel.slowkow.com/)

- R Graph Gallery (https://r-graph-gallery.com/) 

  

